---
title: "ANLY512 Final Project - Quantified Self: Apple Health Dashboard"
author: "Yunwei Sheng"
date: "`r Sys.Date()`"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: flatly
    source_code: embed
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(flexdashboard)
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)
library(htmltools)

data_dir <- "data/processed"
daily <- read_csv(file.path(data_dir, "daily_summary.csv"), show_col_types = FALSE) %>%
  mutate(
    date = as_date(date),
    month = as_date(month),
    weekday = factor(weekday, levels = c(
      "Monday", "Tuesday", "Wednesday", "Thursday",
      "Friday", "Saturday", "Sunday"
    )),
    activity_level = factor(
      case_when(
        steps >= 12000 ~ "High",
        steps >= 6000 ~ "Moderate",
        is.na(steps) ~ NA_character_,
        TRUE ~ "Low"
      ),
      levels = c("Low", "Moderate", "High")
    ),
    workout_prev_day = lag(had_workout),
    high_activity = steps >= quantile(steps, 0.75, na.rm = TRUE)
  )

sleep <- read_csv(file.path(data_dir, "sleep_sessions.csv"), show_col_types = FALSE) %>%
  mutate(wake_date = as_date(wake_date))

workouts <- read_csv(file.path(data_dir, "workouts.csv"), show_col_types = FALSE) %>%
  mutate(
    date = as_date(date),
    month = as_date(month),
    weekday = factor(weekday, levels = levels(daily$weekday))
  )

date_min <- min(daily$date, na.rm = TRUE)
date_max <- max(daily$date, na.rm = TRUE)

kpi_steps <- round(mean(daily$steps, na.rm = TRUE))
kpi_sleep <- round(mean(daily$sleep_hours, na.rm = TRUE), 1)
kpi_workouts <- nrow(workouts)
kpi_hrv <- round(mean(daily$hrv_ms, na.rm = TRUE), 1)

weekday_activity <- daily %>%
  filter(!is.na(steps)) %>%
  group_by(weekday) %>%
  summarise(
    avg_steps = mean(steps),
    avg_sleep = mean(sleep_hours, na.rm = TRUE),
    avg_stand = mean(stand_hours, na.rm = TRUE),
    .groups = "drop"
  )

weekday_sleep <- daily %>%
  filter(!is.na(sleep_hours)) %>%
  group_by(weekday) %>%
  summarise(
    avg_sleep = mean(sleep_hours),
    med_sleep = median(sleep_hours),
    sleep_iqr = IQR(sleep_hours),
    .groups = "drop"
  )

sleep_weekend <- daily %>%
  filter(!is.na(sleep_hours)) %>%
  mutate(period = if_else(weekday %in% c("Saturday", "Sunday"), "Weekend", "Weekday")) %>%
  group_by(period) %>%
  summarise(avg_sleep = mean(sleep_hours), .groups = "drop")

sleep_weekend_diff <- sleep_weekend$avg_sleep[sleep_weekend$period == "Weekend"] -
  sleep_weekend$avg_sleep[sleep_weekend$period == "Weekday"]

sleep_activity_cor <- daily %>%
  filter(!is.na(steps), !is.na(sleep_hours)) %>%
  summarise(r = cor(steps, sleep_hours)) %>%
  pull(r)

activity_sleep <- daily %>%
  filter(!is.na(sleep_hours), !is.na(activity_level)) %>%
  group_by(activity_level) %>%
  summarise(avg_sleep = mean(sleep_hours), .groups = "drop")

very_high_step_sleep <- daily %>%
  filter(!is.na(sleep_hours), steps > 30000) %>%
  summarise(avg_sleep = mean(sleep_hours), days = n())

workout_sleep <- daily %>%
  filter(!is.na(sleep_hours), !is.na(had_workout)) %>%
  group_by(had_workout) %>%
  summarise(avg_sleep = mean(sleep_hours), .groups = "drop")

workout_next_sleep <- daily %>%
  filter(!is.na(sleep_hours), !is.na(workout_prev_day)) %>%
  group_by(workout_prev_day) %>%
  summarise(avg_sleep = mean(sleep_hours), .groups = "drop")

recovery_high <- daily %>%
  filter(!is.na(high_activity), !is.na(hrv_ms)) %>%
  group_by(high_activity) %>%
  summarise(avg_hrv = mean(hrv_ms), avg_rhr = mean(resting_heart_rate, na.rm = TRUE), .groups = "drop")

recovery_step_hrv_cor <- cor(daily$steps, daily$hrv_ms, use = "complete.obs")
recovery_step_rhr_cor <- cor(daily$steps, daily$resting_heart_rate, use = "complete.obs")

roll_peak_steps <- daily %>%
  filter(!is.na(steps_roll7)) %>%
  slice_max(steps_roll7, n = 1, with_ties = FALSE)

roll_min_sleep <- daily %>%
  filter(!is.na(sleep_roll7)) %>%
  slice_min(sleep_roll7, n = 1, with_ties = FALSE)

monthly_trend <- daily %>%
  group_by(month) %>%
  summarise(
    avg_steps = mean(steps, na.rm = TRUE),
    avg_sleep = mean(sleep_hours, na.rm = TRUE),
    avg_energy = mean(active_energy_kcal, na.rm = TRUE),
    .groups = "drop"
  )

month_start <- monthly_trend %>%
  slice_min(month, n = 1, with_ties = FALSE)

month_end <- monthly_trend %>%
  slice_max(month, n = 1, with_ties = FALSE)

workout_types <- workouts %>%
  count(workout_type, sort = TRUE)

workout_monthly <- workouts %>%
  count(month, name = "workouts")

monthly_step_drop <- monthly_trend %>%
  arrange(month) %>%
  mutate(step_change = avg_steps - lag(avg_steps)) %>%
  filter(step_change == min(step_change, na.rm = TRUE))

theme_dash <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```

Overview {data-icon="fa-user"}
=====================================

```{r overview-css, results='asis'}
cat(as.character(tags$style(HTML("
  .navbar-brand,
  .navbar-brand:hover,
  .navbar-brand:focus {
    pointer-events: none;
    cursor: default;
    color: #ffffff !important;
    background-color: transparent !important;
    text-decoration: none;
  }

  .dashboard-page#overview .overview-rq {
    font-size: 16px;
    line-height: 1.65;
  }
"))))
```

Row {data-height=420}
-------------------------------------

### Project Summary {data-width=600}

This dashboard explores **one year of personal Apple Health data** from my Apple Watch and iPhone (June 2025 – June 2026), focusing on how **movement, sleep, and recovery** interact.

- **Data collection:** iPhone Health app export (`export.xml`).
- **Processing:** `scripts/parse_apple_health.R` → three CSV files in `data/processed/`.
- **Tools:** R, `tidyverse`, `lubridate`, `flexdashboard`, `plotly`, `scales`.
- **Approach:** Daily aggregation plus interactive charts to compare habits, workout nights, and weekday patterns.

### Key Metrics (Last 12 Months) {data-width=400}

```{r}
tags$div(
  style = "display:grid; grid-template-columns:1fr 1fr; gap:8px;",
  tags$div(class = "alert alert-info", tags$h4("Avg Daily Steps"), tags$h2(format(kpi_steps, big.mark = ","))),
  tags$div(class = "alert alert-success", tags$h4("Avg Sleep (hrs)"), tags$h2(kpi_sleep)),
  tags$div(class = "alert alert-warning", tags$h4("Workout Sessions"), tags$h2(kpi_workouts)),
  tags$div(
    class = "alert",
    style = "background-color:#ff7f6e; border-color:#ff6b57; color:#ffffff;",
    tags$h4("Avg HRV (ms)"),
    tags$h2(kpi_hrv)
  )
)
```

Row {data-height=300}
-------------------------------------

### Research Questions

```{r}
tags$div(
  class = "overview-rq",
  tags$ol(
    tags$li(HTML("<strong>Sleep vs activity:</strong> Do higher step counts or active energy days coincide with longer sleep?")),
    tags$li(HTML("<strong>Workout lag:</strong> On nights after workouts, is sleep duration different from rest days?")),
    tags$li(HTML("<strong>Weekly rhythm:</strong> Which weekdays are most active, and how does sleep/stand time vary?")),
    tags$li(HTML("<strong>Recovery:</strong> After high-activity days, do HRV and resting heart rate suggest adequate recovery?")),
    tags$li(HTML("<strong>Long-term trends:</strong> Are activity, sleep, and energy improving or declining over 12 months?")),
    tags$li(HTML("<strong>Workout profile:</strong> What workout types dominate, and how often do I train each month?"))
  )
)
```

Activity {data-icon="fa-walking"}
=====================================

Row {data-height=600}
-------------------------------------

### Daily Steps & Active Energy (Interactive) {data-width=650}

```{r}
activity_long <- daily %>%
  select(date, steps, active_energy_kcal, steps_roll7, active_energy_roll7) %>%
  pivot_longer(
    c(steps, active_energy_kcal),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(
    metric = factor(
      metric,
      levels = c("steps", "active_energy_kcal"),
      labels = c("Steps", "Active Energy (kcal)")
    ),
    roll_value = if_else(metric == "Steps", steps_roll7, active_energy_roll7)
  ) %>%
  filter(!is.na(value))

p_activity <- activity_long %>%
  ggplot(aes(date, value, color = metric, group = metric)) +
  geom_line(linewidth = 0.4, alpha = 0.5) +
  geom_line(
    aes(y = roll_value),
    linewidth = 0.9,
    na.rm = TRUE
  ) +
  scale_color_manual(values = c("Steps" = "#2166ac", "Active Energy (kcal)" = "#f46d43")) +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~metric, scales = "free_y", ncol = 1) +
  labs(title = "Daily Activity Over the Past Year", x = NULL, y = NULL, color = NULL) +
  theme_dash +
  theme(strip.text = element_text(face = "bold"), legend.position = "none")

ggplotly(p_activity, tooltip = c("x", "y", "colour")) %>%
  layout(height = 420, hovermode = "x")
```

### Analysis {data-width=350}

**How to read:** The top panel shows daily steps and the bottom panel shows active energy in kcal. Thin lines are daily values; thicker lines are 7-day rolling averages that smooth out daily noise. The two panels use separate y-axes, so compare patterns over time rather than the height of the two lines. Hover to see the exact date and value.

**Answer (Q5 – daily trend):** Steps and energy both go up and down from week to week. The 7-day step average is ~**`r comma(round(head(na.omit(daily$steps_roll7), 1)))`** at the start and ~**`r comma(round(tail(na.omit(daily$steps_roll7), 1)))`** at the end. Several days above 40k steps show up in late December (travel). Steps and energy move together on most days (r = **`r round(cor(daily$steps, daily$active_energy_kcal, use = "complete.obs"), 2)`**).

Row {data-height=460}
-------------------------------------

### Monthly Average Steps & Active Energy {data-width=650}

```{r}
energy_coef <- max(monthly_trend$avg_steps, na.rm = TRUE) / max(monthly_trend$avg_energy, na.rm = TRUE)

monthly_long <- monthly_trend %>%
  arrange(month) %>%
  mutate(month_lab = factor(format(month, "%b %Y"), levels = unique(format(month, "%b %Y")))) %>%
  pivot_longer(c(avg_steps, avg_energy), names_to = "metric", values_to = "display_value") %>%
  mutate(
    metric = factor(
      recode(metric, avg_steps = "Steps", avg_energy = "Active Energy (kcal)"),
      levels = c("Steps", "Active Energy (kcal)")
    ),
    plot_value = if_else(
      metric == "Active Energy (kcal)",
      display_value * energy_coef,
      display_value
    ),
    hover = paste0(
      month_lab, "<br>", metric, ": ",
      if_else(metric == "Steps", comma(round(display_value)), paste0(comma(round(display_value)), " kcal"))
    )
  )

p_month_dual <- ggplot(monthly_long, aes(month_lab, plot_value, fill = metric, text = hover)) +
  geom_col(position = position_dodge2(width = 0.9, padding = 0), width = 0.45) +
  scale_fill_manual(values = c("Steps" = "#4393c3", "Active Energy (kcal)" = "#f46d43")) +
  scale_y_continuous(
    name = "Steps",
    labels = label_comma(),
    sec.axis = sec_axis(~ . / energy_coef, name = "Active Energy (kcal)", labels = label_comma())
  ) +
  labs(title = "Monthly Average Steps & Active Energy", x = NULL, fill = NULL) +
  theme_dash +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplotly(p_month_dual, tooltip = "text") %>%
  layout(height = 380, legend = list(orientation = "h", y = -0.24))
```

### Analysis {data-width=350}

**How to read:** Each month has two bars: blue for average daily steps and orange for average active energy. Steps use the left y-axis; active energy uses the right y-axis. Use the bars to compare month-to-month patterns within each metric, and hover for exact values.

**Answer (Q5 – monthly):** **`r format(monthly_trend$month[which.max(monthly_trend$avg_steps)], "%B %Y")`** has the highest average steps (**`r comma(round(max(monthly_trend$avg_steps)))`**/day) — I was traveling and walked a lot. **`r format(monthly_trend$month[which.min(monthly_trend$avg_steps)], "%B %Y")`** has the lowest (**`r comma(round(min(monthly_trend$avg_steps)))`**/day) — cold weather in NYC kept me inside more. That is a **`r round((max(monthly_trend$avg_steps) - min(monthly_trend$avg_steps)) / max(monthly_trend$avg_steps) * 100)`**% drop from Dec to Jan (**`r comma(round(-monthly_step_drop$step_change))`** steps). Energy follows the same pattern but less strongly (r = **`r round(cor(monthly_trend$avg_steps, monthly_trend$avg_energy, use = "complete.obs"), 2)`**). Activity picks up again in spring but stays below December.

Row {data-height=460}
-------------------------------------

### Stand Hours vs Steps {data-width=650}

```{r}
p_stand <- daily %>%
  filter(!is.na(steps), !is.na(stand_hours)) %>%
  ggplot(aes(steps, stand_hours)) +
  geom_point(alpha = 0.35, color = "#542788") +
  geom_smooth(method = "lm", se = TRUE, color = "#807dba") +
  scale_x_continuous(labels = label_comma()) +
  labs(title = "Stand Hours vs Steps", x = "Steps", y = "Stand hours") +
  theme_dash

ggplotly(p_stand) %>%
  layout(height = 380)
```

### Analysis {data-width=350}

**How to read:** Each dot is one day, with steps on the x-axis and stand hours on the y-axis. The fitted line shows the overall relationship between the two measures, and the shaded band shows uncertainty around that trend.

**Answer (Q3 – stand vs steps):** Days with more steps tend to have more stand hours (r = **`r round(cor(daily$steps, daily$stand_hours, use = "complete.obs"), 2)`**). The trend line slopes upward. There are still exceptions: some low-step days reach 10+ stand hours, while a few very high-step days have lower stand hours. Some exceptions may come from missing Apple Watch wear time, when steps, stand hours, workouts, or sleep were not fully recorded.

Sleep {data-icon="fa-bed"}
=====================================

Row {data-height=460}
-------------------------------------

### Sleep Duration by Weekday {data-width=650}

```{r}
p_sleep_box <- daily %>%
  filter(!is.na(sleep_hours)) %>%
  ggplot(aes(weekday, sleep_hours)) +
  geom_boxplot(fill = "#9ecae1", color = "#2166ac", outlier.alpha = 0.3) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "#08306b") +
  labs(title = "Sleep Duration by Day of Week", x = NULL, y = "Hours asleep") +
  theme_dash

ggplotly(p_sleep_box) %>%
  layout(height = 380)
```

### Analysis {data-width=350}

**How to read:** Each box is one weekday. The box covers the middle 50% of nights, the line inside the box is the median, and dots are unusually short or long nights. The dark blue diamond marks the mean.

**Answer (Q3 – weekly rhythm):**

- **Lowest sleep:** **`r weekday_sleep$weekday[which.min(weekday_sleep$avg_sleep)]`** has the lowest mean (**`r round(min(weekday_sleep$avg_sleep), 1)`** hrs) and sits below the other boxes. It also has short outliers near 2–3 hrs.
- **Highest sleep:** **`r weekday_sleep$weekday[which.max(weekday_sleep$avg_sleep)]`** has the highest mean (**`r round(max(weekday_sleep$avg_sleep), 1)`** hrs).
- **Overall pattern:** Most days cluster around 7 hrs, but the gap between the highest and lowest weekday mean is **`r round(max(weekday_sleep$avg_sleep) - min(weekday_sleep$avg_sleep), 1)`** hrs.
- **Most variable day:** **`r weekday_sleep$weekday[which.max(weekday_sleep$sleep_iqr)]`** has the widest box, so sleep varies more night to night.
- **Weekend effect:** Weekend nights average **`r round(sleep_weekend$avg_sleep[sleep_weekend$period == "Weekend"], 1)`** hrs vs **`r round(sleep_weekend$avg_sleep[sleep_weekend$period == "Weekday"], 1)`** hrs on weekdays (**+`r round(sleep_weekend_diff * 60)`** min).

Row {data-height=460}
-------------------------------------

### Sleep vs Daily Steps {data-width=650}

```{r}
p_sleep_scatter <- daily %>%
  filter(!is.na(steps), !is.na(sleep_hours)) %>%
  ggplot(aes(steps, sleep_hours, color = activity_level)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "gray30", inherit.aes = FALSE, aes(x = steps, y = sleep_hours)) +
  scale_x_continuous(labels = label_comma()) +
  scale_color_manual(values = c(Low = "#91bfdb", Moderate = "#fc8d59", High = "#d73027"), na.value = "gray70") +
  labs(
    title = "Sleep vs Daily Steps",
    x = "Daily steps",
    y = "Sleep (hours)",
    color = NULL
  ) +
  theme_dash

ggplotly(p_sleep_scatter) %>%
  layout(
    height = 380,
    margin = list(b = 70),
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.25)
  )
```

### Analysis {data-width=350}

**How to read:** Each point is one day, with steps on the x-axis and sleep hours on the y-axis. The gray line shows the overall linear trend, with the shaded band showing uncertainty.

Activity levels are defined by daily steps:

- **Low:** under 6,000 steps
- **Moderate:** 6,000–11,999 steps
- **High:** 12,000+ steps

**Answer (Q1 – sleep vs activity):** As a continuous measure, daily steps and sleep have a weak negative relationship (r = **`r round(sleep_activity_cor, 2)`**), so the trend line slopes slightly downward. By activity level, low-activity days average **`r round(activity_sleep$avg_sleep[activity_sleep$activity_level == "Low"], 1)`** hrs of sleep, compared with **`r round(activity_sleep$avg_sleep[activity_sleep$activity_level == "High"], 1)`** hrs on high-activity days. Very high-step days above 30k average **`r round(very_high_step_sleep$avg_sleep, 1)`** hrs. The points are still widely spread, so activity level helps summarize the pattern, but it does not explain sleep duration by itself.

Workouts {data-icon="fa-dumbbell"}
=====================================

Row {data-height=460}
-------------------------------------

### Workout Types (Last 12 Months) {data-width=650}

```{r}
p_workout_types <- workout_types %>%
  mutate(workout_type = fct_reorder(workout_type, n)) %>%
  ggplot(aes(n, workout_type, text = paste0(workout_type, ": ", n, " sessions"))) +
  geom_col(fill = "#1b7837") +
  labs(title = "Workout Types (Last 12 Months)", x = "Sessions", y = NULL) +
  theme_dash

ggplotly(p_workout_types, tooltip = "text") %>%
  layout(height = 380)
```

### Analysis {data-width=350}

**How to read:** Each horizontal bar is one workout type, and the bar length shows how many sessions were logged. The chart is sorted from most frequent to least frequent, so the longest bars show the activities that make up most of the workout history.

**Answer (Q6 – workout profile):** **`r workout_types$workout_type[1]`** is the most frequent activity (**`r workout_types$n[1]`** sessions), followed by **`r workout_types$workout_type[2]`** (**`r workout_types$n[2]`**) and **`r workout_types$workout_type[3]`** (**`r workout_types$n[3]`**). These top three types account for **`r sum(workout_types$n[1:3])`** of **`r kpi_workouts`** workouts, so the workout profile is concentrated rather than evenly spread across activities.

Row {data-height=460}
-------------------------------------

### Monthly Workout Count {data-width=650}

```{r}
p_workout_month <- workout_monthly %>%
  ggplot(aes(month, workouts, text = paste0(format(month, "%b %Y"), "<br>Workouts: ", workouts))) +
  geom_col(fill = "#762a83", width = 20) +
  labs(title = "Monthly Workout Count", x = NULL, y = "Sessions") +
  theme_dash

ggplotly(p_workout_month, tooltip = "text") %>%
  layout(height = 380)
```

### Analysis {data-width=350}

**How to read:** Each bar shows the number of workouts logged in one month. Taller bars mean higher training frequency. Use this chart to compare monthly consistency and identify unusually high or low workout months.

**Answer (Q6 – frequency):** Monthly volume peaks in **`r format(workout_monthly$month[which.max(workout_monthly$workouts)], "%B %Y")`** with **`r max(workout_monthly$workouts)`** sessions and drops to **`r min(workout_monthly$workouts)`** in **`r format(workout_monthly$month[which.min(workout_monthly$workouts)], "%B %Y")`**. The bars show two active periods, late summer/early fall and spring 2026, with a clear low point in December. December still had high step counts, so much of that activity came from walking while traveling rather than formal logged workouts.

Row {data-height=460}
-------------------------------------

### Average Sleep: Workout vs Rest Days {data-width=650}

```{r}
sleep_same_diff_min <- round(
  (workout_sleep$avg_sleep[workout_sleep$had_workout] -
     workout_sleep$avg_sleep[!workout_sleep$had_workout]) * 60
)
sleep_next_diff_min <- round(
  (workout_next_sleep$avg_sleep[workout_next_sleep$workout_prev_day] -
     workout_next_sleep$avg_sleep[!workout_next_sleep$workout_prev_day]) * 60
)

sleep_workout_plot <- tibble(
  group = factor(
    c("Rest day", "Workout day", "Day after rest", "Day after workout"),
    levels = c("Rest day", "Workout day", "Day after rest", "Day after workout")
  ),
  pair = factor(rep(c("Same night", "Next night"), each = 2), levels = c("Same night", "Next night")),
  avg_sleep = c(
    workout_sleep$avg_sleep[!workout_sleep$had_workout],
    workout_sleep$avg_sleep[workout_sleep$had_workout],
    workout_next_sleep$avg_sleep[!workout_next_sleep$workout_prev_day],
    workout_next_sleep$avg_sleep[workout_next_sleep$workout_prev_day]
  )
) %>%
  group_by(pair) %>%
  mutate(
    diff_min = round((avg_sleep - first(avg_sleep)) * 60),
    top_label = if_else(
      diff_min == 0,
      paste0(round(avg_sleep, 2), " h"),
      paste0(round(avg_sleep, 2), " h  (+", diff_min, " min)")
    )
  ) %>%
  ungroup()

sleep_ylim <- c(
  min(sleep_workout_plot$avg_sleep) - 0.06,
  max(sleep_workout_plot$avg_sleep) + 0.22
)

p_workout_sleep <- sleep_workout_plot %>%
  ggplot(aes(group, avg_sleep, fill = group, text = paste0(
    group, "<br>", round(avg_sleep, 2), " hrs",
    if_else(diff_min == 0, "", paste0("<br>+", diff_min, " min vs rest"))
  ))) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_label(
    aes(label = top_label),
    vjust = 0.25,
    size = 3.4,
    fontface = "bold",
    fill = "white",
    color = "#333333",
    label.size = 0.15,
    label.padding = unit(0.18, "lines"),
    show.legend = FALSE
  ) +
  scale_fill_manual(values = c(
    "Rest day" = "#66c2a5",
    "Workout day" = "#fc8d62",
    "Day after rest" = "#8da0cb",
    "Day after workout" = "#e78ac3"
  )) +
  facet_wrap(~pair, scales = "free_x") +
  coord_cartesian(ylim = sleep_ylim) +
  labs(
    title = "Average Sleep: Workout vs Rest Days",
    subtitle = "Paired panels with zoomed y-axis; labels show hours and minute gain vs rest baseline",
    x = NULL,
    y = "Hours asleep"
  ) +
  theme_dash +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    strip.text = element_text(face = "bold")
  )

ggplotly(p_workout_sleep, tooltip = "text") %>%
  layout(height = 430, margin = list(t = 60, b = 55), showlegend = FALSE)
```

### Analysis {data-width=350}

**How to read:** The left panel compares sleep on rest days vs workout days. The right panel compares sleep after rest days vs sleep after workout days. The y-axis is zoomed because the differences are small; labels show average hours and the minute difference from the rest baseline.

**Answer (Q2 – workout lag):** Workout days average **`r sleep_same_diff_min`** more minutes of sleep on the same night and **`r sleep_next_diff_min`** more minutes the following night. Both differences are positive but small. This suggests workouts are not associated with shorter sleep in this dataset, but the size of the difference should not be overinterpreted.

Recovery {data-icon="fa-heartbeat"}
=====================================

Row {data-height=460}
-------------------------------------

### HRV vs Daily Steps {data-width=650}

```{r}
p_recovery_scatter <- daily %>%
  filter(!is.na(steps), !is.na(hrv_ms)) %>%
  ggplot(aes(steps, hrv_ms, color = resting_heart_rate, text = paste0(
    format(date, "%b %d"), "<br>Steps: ", comma(steps),
    "<br>HRV: ", round(hrv_ms, 1), " ms"
  ))) +
  geom_point(alpha = 0.55, size = 2) +
  geom_smooth(aes(group = 1), method = "lm", se = TRUE, color = "gray30") +
  scale_x_continuous(labels = label_comma()) +
  scale_color_gradient(low = "#2166ac", high = "#b2182b", na.value = "gray80") +
  labs(title = "HRV vs Daily Steps", x = "Steps", y = "HRV (ms)", color = "Resting HR") +
  theme_dash

ggplotly(p_recovery_scatter, tooltip = "text") %>%
  layout(height = 380)
```

### Analysis {data-width=350}

**How to read:** Each point is one day, with steps on the x-axis and HRV on the y-axis. HRV means heart rate variability: higher HRV usually reflects better recovery or lower stress, while lower HRV can reflect fatigue, stress, or incomplete recovery. Color shows resting heart rate: blue is lower and red is higher. The gray line shows the overall relationship between steps and HRV.

**Answer (Q4 – recovery):** The trend is negative: higher-step days tend to have lower HRV (r = **`r round(recovery_step_hrv_cor, 2)`**). High-activity days average **`r round(recovery_high$avg_hrv[recovery_high$high_activity], 1)`** ms HRV vs **`r round(recovery_high$avg_hrv[!recovery_high$high_activity], 1)`** ms on lower-activity days. Resting heart rate is also higher on high-activity days (**`r round(recovery_high$avg_rhr[recovery_high$high_activity], 1)`** vs **`r round(recovery_high$avg_rhr[!recovery_high$high_activity], 1)`** bpm; steps vs RHR r = **`r round(recovery_step_rhr_cor, 2)`**). In plain terms, high-step days often combine more physical load with a recovery signal that looks more strained: lower HRV and higher resting heart rate.

Row {data-height=540}
-------------------------------------

### 7-Day Rolling Averages {data-width=650}

```{r}
roll_long <- daily %>%
  select(date, steps_roll7, sleep_roll7, active_energy_roll7) %>%
  pivot_longer(-date, names_to = "metric", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(
    metric = factor(
      recode(
        metric,
        steps_roll7 = "Steps (7-day avg)",
        sleep_roll7 = "Sleep (7-day avg, hrs)",
        active_energy_roll7 = "Energy (7-day avg, kcal)"
      ),
      levels = c(
        "Steps (7-day avg)",
        "Energy (7-day avg, kcal)",
        "Sleep (7-day avg, hrs)"
      )
    )
  )

p_roll <- roll_long %>%
  ggplot(aes(date, value, color = metric, group = metric)) +
  geom_line(linewidth = 0.75) +
  scale_color_manual(values = c(
    "Steps (7-day avg)" = "#4393c3",
    "Energy (7-day avg, kcal)" = "#f4a582",
    "Sleep (7-day avg, hrs)" = "#807dba"
  )) +
  scale_y_continuous(labels = label_comma()) +
  facet_wrap(~metric, scales = "free_y", ncol = 1) +
  labs(title = "7-Day Rolling Averages: Activity, Sleep & Energy", x = NULL, y = NULL) +
  theme_dash +
  theme(legend.position = "none", strip.text = element_text(face = "bold"))

ggplotly(p_roll, tooltip = c("x", "y", "colour")) %>%
  layout(height = 500, hovermode = "x")
```

### Analysis {data-width=350}

**How to read:** Each panel shows a 7-day rolling average, which smooths daily noise. Steps, active energy, and sleep use separate y-axes, so compare each line's direction over time rather than comparing panel heights.

**Answer (Q5 – long-term trends):** The largest activity spike appears in late December, when the 7-day step average peaks around **`r comma(round(roll_peak_steps$steps_roll7))`** steps. Sleep does not rise with that spike; its lowest 7-day average is around **`r round(roll_min_sleep$sleep_roll7, 1)`** hrs near the same period. Because this was a travel period, time-zone changes and inconsistent watch wear may have affected sleep recording. Comparing the same month across years, **`r format(month_end$month, "%b %Y")`** is higher than **`r format(month_start$month, "%b %Y")`** in both average steps (**`r comma(round(month_end$avg_steps))`** vs **`r comma(round(month_start$avg_steps))`**) and average sleep (**`r round(month_end$avg_sleep, 1)`** vs **`r round(month_start$avg_sleep, 1)`** hrs).

Insights {data-icon="fa-lightbulb"}
=====================================

Row {data-height=500}
-------------------------------------

### Weekly Profile Heatmap {data-width=650}

```{r}
week_heat <- weekday_activity %>%
  select(weekday, avg_steps, avg_sleep, avg_stand) %>%
  pivot_longer(-weekday, names_to = "metric", values_to = "value") %>%
  mutate(
    metric = factor(
      recode(metric,
        avg_steps = "Steps",
        avg_sleep = "Sleep (hrs)",
        avg_stand = "Stand hrs"
      ),
      levels = c("Steps", "Sleep (hrs)", "Stand hrs")
    )
  ) %>%
  group_by(metric) %>%
  mutate(
    value_scaled = if (max(value) == min(value)) {
      0.5
    } else {
      (value - min(value)) / (max(value) - min(value))
    }
  ) %>%
  ungroup()

p_heat <- ggplot(week_heat, aes(metric, weekday, fill = value_scaled, text = paste0(
    weekday, " · ", metric, "<br>",
    if_else(metric == "Steps", paste0(comma(round(value)), " steps"),
      if_else(metric == "Sleep (hrs)", paste0(round(value, 1), " hrs"),
        paste0(round(value, 1), " stand hrs")
      )
    )
  ))) +
  geom_tile(color = "white", linewidth = 0.8) +
  scale_fill_gradient(low = "#deebf7", high = "#2166ac") +
  labs(
    title = "Weekday Patterns (Normalized)",
    subtitle = "Each column scaled separately so steps, sleep, and stand hours are comparable by weekday",
    x = NULL,
    y = NULL,
    fill = "Relative level"
  ) +
  theme_dash

ggplotly(p_heat, tooltip = "text") %>%
  layout(height = 440, margin = list(b = 40))
```

### Analysis {data-width=350}

**How to read:** Each column is scaled separately from 0 to 1, so the darkest cell means "highest weekday for this metric," not the largest raw value across all metrics. Use the heatmap to compare weekday patterns within Steps, Sleep, and Stand hours.

**Answer (Q3 – weekly pattern):** **`r weekday_activity$weekday[which.max(weekday_activity$avg_steps)]`** is darkest for Steps and also high for Stand hours, so it is the most active weekday overall. **`r weekday_activity$weekday[which.max(weekday_activity$avg_sleep)]`** is darkest for Sleep, while **`r weekday_activity$weekday[which.min(weekday_activity$avg_sleep)]`** is the lightest sleep day. The chart shows that activity and sleep do not peak on the same weekday.

Row {data-height=360}
-------------------------------------

### Final Summary

**Overall summary:** This dashboard uses one year of my own Apple Health data to compare movement, sleep, workouts, and recovery. The main pattern is that activity, sleep, and recovery move together sometimes, but not perfectly. Higher movement usually raises energy and stand hours, but it does not automatically mean longer sleep or better recovery.

**Research question answers:** Step count and sleep have a weak negative relationship (r = **`r round(sleep_activity_cor, 2)`**), so higher activity days are not clearly linked with longer sleep. Workouts are associated with slightly more sleep: **`r round(workout_sleep$avg_sleep[workout_sleep$had_workout], 2)`** hrs on workout nights and **`r round(workout_next_sleep$avg_sleep[workout_next_sleep$workout_prev_day], 2)`** hrs after workout days, but the differences are small. Weekly rhythm is visible in the heatmap: **`r weekday_activity$weekday[which.max(weekday_activity$avg_steps)]`** has the highest average steps and stand hours, while **`r weekday_activity$weekday[which.max(weekday_activity$avg_sleep)]`** has the highest sleep. Recovery is the main caution: high-activity days have lower HRV (**`r round(recovery_high$avg_hrv[recovery_high$high_activity], 1)`** vs **`r round(recovery_high$avg_hrv[!recovery_high$high_activity], 1)`** ms) and higher resting HR (**`r round(recovery_high$avg_rhr[recovery_high$high_activity], 1)`** vs **`r round(recovery_high$avg_rhr[!recovery_high$high_activity], 1)`** bpm).

**Long-term pattern:** December 2025 is the activity peak, largely because I was traveling and walking more. At the same time, December has very few formal workout sessions, which suggests much of the activity came from travel walking rather than intentional workouts. January 2026 is the low point, when cold weather in NYC kept me indoors more. Activity rebounds in spring, but it does not fully return to the December peak.

**Data limitation:** These patterns depend on Apple Watch and iPhone records. Some outliers may reflect times when I was not wearing the watch continuously, so certain workouts, stand hours, steps, or sleep sessions may be missing or incomplete. The dashboard also does not control for outside factors such as caffeine, travel/jet lag, stress, or illness, which can affect sleep and recovery independently of activity.

**Final takeaway:** My routine is strongest when activity is consistent rather than extreme. Tennis dominates the workout profile (**`r workout_types$workout_type[1]`**, **`r workout_types$n[1]`** sessions), and workouts do not appear to hurt sleep. However, very high activity can coincide with lower HRV, so recovery metrics add context that steps alone cannot show.
